---
title: "Class 3 Project"
format: 
  html:
    embed-resources: true
---

This file serves as a template for the Class 3 in-class project.

```{r}
#| warning: false
library(tidyverse)     # loads the tidyverse tools
library(RPostgres)     # loads the database driver for PostgreSQL
library(connections)   # helps RPostgres work with RStudio
library(keyring)       # access to a local encrypted keychain for passwords

con <- connection_open(RPostgres::Postgres(),
          dbname = "syntheticMGUH",
          host = "34.85.177.140",
          user = "hids502_student",
          password = key_get(service = "syntheticmguh", 
                             username = "hids502_student"))
```

# Question 1

How many patients do we have in the database?

```{sql connection=con}
SELECT COUNT(*) AS patient_count
FROM patients 
```


How many providers and payers are represented in the data set?

```{sql connection=con}
-- Your code goes here
```


# Question 2

What is the mean amount spent per patient?

What does this look like broken down by year?

# Question 3

What is the mean amount spent for Medicare patients?

What does this look like broken down by  year?

# Question 4

Using a common table expression, what are the min, max, and mean number of patients per provider?

Using the raw data in R, plot this as a boxplot as well.

Here's an example of a boxplot (note the additional use of the `y` axis that you probably don't need here.)

```{sql connection=con, output.var="expenses_df"}
-- Here we use `output.var` to assign the results of the query to a variable
-- rather than printing!
SELECT id, race, healthcare_expenses FROM patients
```

```{r}
ggplot(expenses_df) + 
  geom_boxplot(aes(x = healthcare_expenses, y = race))
```

# Question 5

What is the min/max and mean number of encounters per patient in 2022? 

Plot the distribution as a histogram – this will require you to use R

Here's an example of a boxplot

```{r}
ggplot(expenses_df) + 
  geom_histogram(aes(x = healthcare_expenses))
```
# Question 6 (Optional)

What time frame does the database cover? In other words, when does the data start and when does it end?

Show the volume of visits over time (as a table or a plot)

```{sql connection=con}
-- Your code goes here
```
